---
title: "Assignment 3"
author: "Anuraag Vasal"
date: "10/18/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(dplyr)
library(fastDummies)
library(caret)
library(class)
library(melt)
library(reshape)
library(reshape2)
library(ggplot2)
library(ISLR)
library(naivebayes)
library(e1071)

setwd("C:/Users/anura/Desktop/Machine Learning/Assignment 3")
Bank<- read.csv("C:/Users/anura/Desktop/Machine Learning/Assignment 3/UniversalBank.csv")

##Transform Data
Bank$PersonalLoan <- as.factor(Bank$Personal.Loan)
Bank$Online <- as.factor(Bank$Online)
Bank$CreditCard <- as.factor(Bank$CreditCard)
View(Bank)
set.seed(15)

#Data Partition
Train_Index <- createDataPartition(Bank$Personal.Loan, p=0.6, list=FALSE)
Train <-Bank[Train_Index,]
Valid <- Bank[-Train_Index,]
#{A}
Melt_Train <- melt(Train,id=c("CreditCard","Personal.Loan"),variable= "Online")
cast_Train <- dcast(Melt_Train,CreditCard+Personal.Loan~Online)
cast_Train <-cast_Train[c(1,2,14)]
cast_Train
View(cast_Train)
```
#{B}
The probability of the customer accepting the loan is 90/888 =0.10135

#{C}
```{r}
#pivot table1
Melt_Train1 <- melt(Train,id=c("Personal.Loan"),variable= "Online")
cast_Train1 <- dcast(Melt_Train1,Personal.Loan~Online)
cast_Train1 <-cast_Train1[c(1,13)]
cast_Train1
#pivot table2
Melt_Train2 <- melt(Train,id=c("CreditCard"),variable= "Online")
cast_Train2 <- dcast(Melt_Train2,CreditCard ~ Online)
cast_Train2 <-cast_Train2[c(1,14)]
cast_Train2
```
#{D}Compute the following quantities [P(A | B) means “the probability of A given B. 
```{r}
Train_Data <- Train[c(13,10,14)]
table(Train_Data[,c(3,2)])
table(Train_Data[,c(1,2)])
table(Train_Data[,c(2)])
```
i)   P(CC=1 | PL=1)  = 90/(798+90)= 0.10135
ii)  P(OL=1 | PL=1)  = 180/(180+119)=0.60200
iii) P(PL=1)         = (209+90)/3000=0.09966
iv)  P(CC=1 | PL=0)  = 798/(798+1903)=0.29544
v)   P(OL=1 | PL=0)  = 1607/(1607+1094)=0.59496
vi)  P(PL=0)         = (1903+798)/3000=0.90033

#{E}
Use the quantities computed above to compute the naive Bayes probability P(Loan = 1 | CC = 1, Online = 1).
 =(0.10135*0.60266*0.09966)/((0.10135*0.60266*0.09966)+(0.29544*0.59496*0.90033))
 = (0.00608052)/(0.00608052+0.15825548)
 = 0.03700053
 
#{F} Compare this value with the one obtained from the pivot table in (B). Which is a more accurate estimate?
ANS:- Part B = 0.10135  and Part E = 0.03700053
      by comparing both the values part B is more accurate.
#{G}
```{r}
train.df <- Bank[Train_Index, ]
test.df <- Bank[ -Train_Index,]
train <- Bank[Train_Index, ]
test <- Bank[-Train_Index,]
nb_train = train.df[,c(10,13:14)]
nb_test = test.df[,c(10,13:14)]
naivebayes = naiveBayes(Personal.Loan~.,data=nb_train)
naivebayes

```
