---
title: "Final Project"
author: "Anuraag Vasal"
date: "12/10/2021"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(dplyr)
library(ggplot2)
library(data.table)
library(ggcorrplot)
library(pastecs)
library(corrplot)
```

```{r}
setwd("C:/Users/anura/Desktop/Machine Learning Final")
list.files()
raw.data <- read.csv("Retail Marketing.csv")
str(raw.data)

#There are 10 variables and 1000 records (before cleaning the data)
```

#PART I : Cleaning and orginazing the data 
```{r}
str(raw.data)
table(is.na(raw.data$Age))
table(is.na(raw.data$Gender))
table(is.na(raw.data$OwnHome))
table(is.na(raw.data$Married))
table(is.na(raw.data$Location))
table(is.na(raw.data$Salary))
table(is.na(raw.data$Children))
table(is.na(raw.data$History))
table(is.na(raw.data$Catalogs))
table(is.na(raw.data$AmountSpent))
```
#Replacing NA in History with 'Unknown':
```{r}
raw.data$History <- as.character(raw.data$History)
raw.data$History[is.na(raw.data$History)] <- 'Unknown'
raw.data$History <- factor(raw.data$History)
table((raw.data$History))
```
#Removing the 6 NAs in no amount spent
```{r}
retail.df <- raw.data[!is.na(raw.data$AmountSpent),]
```
#factorizing the variable Children 
```{r}
retail.df$Children <- factor(retail.df$Children)  
View(retail.df %>%
       group_by(Catalogs) %>%
       summarise(mean_of_amount = mean(AmountSpent),numebr_of_appirances = n()))
```
#By looking at the table above variable Catalogs is actually a factor variable where 6 is the 'low_end' products and 24 is 'high_end' products,where as 12 and 16 are the mid_range products.

#Changing the notation to more intuitive notation.
```{r}
retail.df<- (retail.df %>%
       mutate(Catalog = ifelse (Catalogs ==6, 'low_end',
                                (ifelse(Catalogs == 12, "low_midrange", 
                                        (ifelse(Catalogs == 18, "high_midrange", "high_end")))))))
```
#Factorizing the new variable
```{r}
retail.df$Catalog <- as.factor(retail.df$Catalog)

#And removing the old one:
retail.df$Catalogs <- NULL
str(retail.df) 
#We are left with 10 variables, 8 of them are factors and 2 are integers (salary + amount spent)
```
#PART II :Summary and Statistics

#Distribution of each categorical variables
```{r}
lapply( retail.df %>%
          select(c("Age", "Gender", "OwnHome", "Married", "Location", "Children", "History","Catalog"))
        ,table)

ggplot(data = retail.df, aes(x = Salary))+
  geom_histogram(bins = 50, colour = 'white', fill = 'darkblue')+
  scale_x_continuous(breaks = seq(0,150000,25000))+
  scale_y_continuous(breaks = seq(0,70,10))+
  xlab("Salary")+
  ylab("Frequency")+
  ggtitle("Distribution of salaries")+
  geom_vline(xintercept = mean(retail.df$Salary), color = 'red')+
  labs(subtitle  = 'red line represent average salary')
  
```
```{r}
mean_salary_female <- mean(retail.df$Salary[retail.df$Gender =="Female"])
mean_salary_male <- mean(retail.df$Salary[retail.df$Gender =="Male"])


ggplot(data = retail.df, aes(x = Salary))+
  geom_histogram(bins = 50, colour = 'white', fill = 'darkblue')+
  scale_x_continuous(breaks = seq(0,150000,35000))+
  scale_y_continuous(breaks = seq(0,70,10))+
  xlab("Salary")+
  ylab("Frequency")+
  ggtitle("Distribution of salaries by gender")+
  geom_vline(xintercept = mean_salary_female, color = 'pink',size=1.5)+
  geom_vline(xintercept = mean_salary_male, color = 'red', alpha= 0.6)+
  labs(subtitle  = "Red line represents male average salary, and pink line represents female average salary")+
  facet_wrap(~Gender)
```
```{r}
mean_AmountSpent_female <- mean(retail.df$AmountSpent[retail.df$Gender =="Female"])
mean_AmountSpent_male <- mean(retail.df$AmountSpent[retail.df$Gender =="Male"])

ggplot(data = retail.df, aes(x = AmountSpent))+
  geom_histogram(bins = 50, colour = 'white', fill = 'lightgreen')+
  scale_x_continuous()+
  scale_y_continuous()+
  xlab("Amount Spent")+
  ylab("Frequency")+
  ggtitle("Distribution of Amount Spent by gender")+
  labs(subtitle  = "Red line represents average amount spent by male and pink line represents average amount spent by female")+
  facet_wrap(~Gender)+
  geom_vline(xintercept = mean_AmountSpent_female, color = 'pink',size=1.5)+
  geom_vline(xintercept = mean_AmountSpent_male, color = 'red', alpha= 0.6)
```

```{r}
#raw.data <- read.csv("Retail Marketing.csv")
raw.data <- raw.data[!is.na(raw.data$AmountSpent),]
head(raw.data$Age, 10)

cor.data <- raw.data
levels(raw.data$Age)
cor.data$Age <- ifelse(cor.data$Age == 'Young', 0,
                        ifelse(cor.data$Age == 'Middle',1,2))

levels(raw.data$Gender)
cor.data$Gender <- ifelse(cor.data$Gender == "Female", 0 ,1)
levels(raw.data$OwnHome)
cor.data$OwnHome <- ifelse(cor.data$OwnHome == "Rent", 0 ,1)
levels(raw.data$Married)
cor.data$Married <- ifelse(cor.data$Married == "Single", 0 ,1)
levels(raw.data$Location)
cor.data$Location_close <- ifelse(cor.data$Location == "Far", 0 ,1)
cor.data$History<- NULL
cor.data$Location<- NULL
```

```{r}
str(cor.data)
cor.maxtrix<- cor(cor.data, method = "pearson", use = "complete.obs")
corrplot(cor.maxtrix)
```

```{r}
library(ggplot2)
par(mfrow=c(1,1))
barplot(table(raw.data$Age), main="Age", col = "#69b3a2")
barplot(table(raw.data$Gender), main="Gender", col = "#A9A9A9")
barplot(table(raw.data$OwnHome), main="Own Home?", col = "#69b3a2")
barplot(table(raw.data$Married), main="Married", col = "#A9A9A9")
barplot(table(raw.data$Location), main="Location", col = "#69b3a2")
barplot(table(raw.data$Children), main="Children", col = "#A9A9A9")
barplot(table(raw.data$Catalog), main="Catalog", col = "#69b3a2")
par(
  mfrow=c(1,2),
  mar=c(4,4,1,0)
)
hist((raw.data$AmountSpent), xlab="", main="Amount Spent", col = "#69b3a2")
hist((raw.data$Salary), xlab="", ylab="", main="Salary", col = "#A9A9A9", grid=2)
```
```{r}
library(cluster)
library(factoextra)
library(flexclust)
library(fpc)
library(clustertend)
library(ClusterR)
library(data.table)

retail.df <- raw.data[!is.na(raw.data$AmountSpent),]
```

```{r}
clustering.df <- cor.data
dim(clustering.df)[2]
```
#Choosing optimal number of clusters

#Assuming the maximum K to cluster is 10: 
```{r}
k.max <- 10

#we will create a vector of the total within sum of squars, in order to visulize it 
wss <- sapply(1:k.max, function(k){kmeans(clustering.df, k, nstart=50,iter.max = 1000 )$tot.withinss})

options("scipen"=999)
ggplot()+ aes(x = 1:k.max, y = wss) + geom_point() + geom_line()+
  labs(x = "Number of clusters K", y = "Total within-clusters sum of squares")+
  scale_x_continuous(breaks = seq(0,10,1))+
  ggtitle("The Elbow Method")
```
```{r}
fviz_nbclust(clustering.df, FUN = kmeans,method = "wss" ,nstart = 50)
```

#Cannot determine optimal number of clusters, when looking at the Elbow Method, therefore using silhouette score.
```{r}
opt.k.sil<- Optimal_Clusters_KMeans(clustering.df, max_clusters=10, plot_clusters=TRUE, 
                                    criterion="silhouette")
#both 2 and 4 number of clusters generated a high silhouette score of 5.9 
#combining that with the WSS output we can conclude that the optimal number of clusters would be 4. 
```
#Clustering
```{r}
retail.df$History <- NULL
retail.df <- raw.data[!is.na(raw.data$AmountSpent),]
KMC <- kmeans(clustering.df,centers = 4,iter.max = 999, nstart=50) 
retail.clustered <- (cbind(retail.df, cluster= KMC$cluster)) 
```
#Creating a new DF, consisted with the original DF with the cluster number for each observation.
```{r}
table_of_cluster_distribution <- table(retail.clustered$cluster)
#   The result:
#  1   2   3   4 
# 283 157 285 269 
table_of_cluster_distribution
barplot(table_of_cluster_distribution, xlab="Clusters", 
        ylab="# of customers", main="# of customers in each cluster",
        col="#69b3a2")

retail.clustered <- data.table(retail.clustered)
retail.clustered[, avg_AmountSpent_in_cluster := mean(AmountSpent),by=list(cluster)]
retail.clustered[, avg_SalarySpent_in_cluster := mean(Salary),by=list(cluster)]

retail.clustered  <-  retail.clustered[, c("Age", "Gender", "OwnHome", "Married",                   
         "Location", "Children", "Catalogs", "Salary","AmountSpent", 
         "avg_AmountSpent_in_cluster", "avg_SalarySpent_in_cluster", "cluster" )]

cluster_1 <- retail.clustered[retail.clustered$cluster==1,]
cluster_2 <- retail.clustered[retail.clustered$cluster==2,]
cluster_3 <- retail.clustered[retail.clustered$cluster==3,]
cluster_4 <- retail.clustered[retail.clustered$cluster==4,]
```
